AggregateCounts

### 2021-07-14 R/AggregateCounts.Rmd
added culture and strain columns to data and all worked

```{r aggregate input}
aggregateList<-GetAggregates()#this reads the data from Matlab
#remove bad samples and irrelevant conditions
##first remove irrelevant conditions
removable <- bind_rows(aggregateList %>% filter(Species == "Pa" & Condition == "mono"), #Pa in Sa mono culture
                       aggregateList%>% filter(Species == "Sa" & Condition == "pawt"),#Sa in all Pa mono cultures
                       aggregateList%>% filter(Species == "Sa" & Condition == "papqsL"),
                       aggregateList%>% filter(Species == "Sa" & Condition == "paphz"))
aggregateList <- anti_join(aggregateList, removable)
##then, bad samples
aggregateListRemoved<-aggregateList %>% filter(!SampleCode == "mono_d1_03_Pa" & !SampleCode == "mono_d1_03_Sa")
aggregateListRemoved<-aggregateList %>% filter(!SampleCode == "phz_d4_14_Pa" & !SampleCode == "phz_d4_14_Sa")
aggregateListRemoved<-aggregateList %>% filter(!SampleCode == "pqsL_d4_02_Pa" & !SampleCode == "pqsL_d4_02_Sa")
aggregateListRemoved<-aggregateList %>% filter(!SampleCode == "pqsL_d4_04_Pa" & !SampleCode == "pqsL_d4_04_Sa")
aggregateListRemoved<-aggregateList %>% filter(!SampleCode == "pqsL_d4_10_Pa" & !SampleCode == "pqsL_d4_10_Sa")
aggregateListRemoved<-aggregateList %>% filter(!SampleCode == "wt_d4_07_Pa" & !SampleCode == "wt_d4_07_Sa")
aggregateListRemoved<-aggregateList %>% filter(!SampleCode == "papqsL_d4_02_Pa" & !SampleCode == "papqsL_d4_02_Sa")
aggregateListRemoved<-aggregateList %>% filter(!SampleCode == "pawt_d4_05_Pa" & !SampleCode == "pawt_d4_05_Sa")
aggregateListRemoved<-aggregateList %>% filter(!SampleCode == "phz_d4_06_Pa" & !SampleCode == "phz_d4_06_Sa")

#add proper labels for plots
aggregateList$Species <- factor(aggregateList$Species, levels = c("Pa", "Sa"),labels = c('P. aeruginosa','S. aureus'))
aggregateList$Condition <- factor(aggregateList$Condition, levels = c("mono", "pawt","papqsL", "paphz", "wt", "pqsL", "phz"),
                                  labels = c('Sa mono', 'Pa wt mono','pqsL mono','phz mono','wt co', 'pqsL co', 'phz co'))
aggregateList$TimePoint <- factor(aggregateList$TimePoint, levels = c("d1", "d4"),labels = c('Day 1','Day 4'))
#agregate input functions
ReadAggregateFile<-function(file, species, timepoint, condition, replicate, sampleCode){#make a dataframe of the aggregate size list and add experimental parameters
  print(file)
   datos <-read.csv(file, header = F)  
   colnames(datos)<-c('Size') 
   datos<-datos %>% filter(!Size<0.5) #first filter the size
   datos<-datos %>% mutate(Species = species)#add parameters
   datos<-datos %>% mutate(TimePoint = timepoint) 
   datos<-datos %>% mutate(Condition = condition)
   datos<-datos %>% mutate(Replicate = replicate)
   datos<-datos %>% mutate(SampleCode = sampleCode)
   datos$Species<-as.factor(datos$Species)
   datos$Replicate<-as.factor(datos$Replicate)
   datos$Condition<-as.factor(datos$Condition)
   datos$TimePoint<-as.factor(datos$TimePoint)
   return(datos)
}
GetAggregates<-function(){#import all experiments at once
   folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Aggregate lists')
   fileList<-list.files(path = folder, full.names = TRUE)
   growthDf<-data.frame()
   for (aggList in 1:length(fileList)){
     fileName<-str_split(basename(fileList[aggList]), "\\.")[[1]][1]#this selects the base file name and removes the ".csv"
     attrList<-str_split(fileName,"_")[[1]]
     growthDf<-rbind(growthDf,ReadAggregateFile(fileList[aggList],attrList[4],attrList[2],attrList[1],attrList[3], fileName))
     }
   return(growthDf)
}
```

@@

#Get aggregate summary per sample
aggregateListSummaryReplicate<-aggregateList%>% group_by(Species, TimePoint, Condition, Replicate) %>% 
  summarise(Aggs = n(), biomass = sum(Size), meanAggSize = mean(Size), seAggSize = std.error(Size), maxSize = max(Size), minSize = min(Size)) %>%
  ungroup() %>% mutate(RatioMaxTotal = maxSize/biomass )

@@

ggplot()+ 
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= meanAggSize, color = Condition,shape = "MeanSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= maxSize, color = Condition, shape = "MaxSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= biomass, color = Condition, shape = "TotalBiomass"),size = 3, height = 0,width= .3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(.~Species, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per sample')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
  theme_cowplot()+
  panel_border() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facet text
        strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))#facet background

@@

#Get aggregate Summary per condition
aggregateListSummary<-aggregateList %>% group_by(Species, TimePoint, Condition) %>% 
  summarise(Aggs = n(), biomass = sum(Size), meanAggSize = mean(Size), seAggSize = std.error(Size), maxSize = max(Size), minSize = min(Size)) %>%
  ungroup() %>% mutate(RatioMaxTotal = maxSize/biomass ) 
#plot agg summary per condition
#####
ggplot()+ 
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= meanAggSize, color = Condition,shape = "MeanSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= maxSize, color = Condition, shape = "MaxSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= biomass, color = Condition, shape = "TotalBiomass"),size = 3, height = 0,width= .3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(. ~ Species, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per condition')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5),strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))
#####
```

@@

#There are two ways to normalize the Bin Frequncy and Bin Biomass, per condition and per sample, so I'll do both
#First per replicate:
HistReplicate<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% group_by(Species,Condition,TimePoint, Replicate)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>% group_by(Species,Condition,TimePoint, Replicate, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

#This is weird, but I want to have the histogram above with all the data, and the one below with only the bins
HistReplicateSummary<-HistReplicate %>% group_by(Species,Condition,TimePoint, Replicate, LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint, Replicate) %>% mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))

#Now per condition. The difference is int the "group_by" part
HistCondition<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% group_by(Species,Condition,TimePoint)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>%
  group_by(Species,Condition,TimePoint,LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

#Summary data frame
HistConditionSummary<-HistCondition %>%
  group_by(Species,Condition,TimePoint,LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>% 
  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint) %>% 
  mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))

#Plot Normalized Frequency
#this isn't useful as the normalized Frequency changes with replicate, so it's better to pool all samples together
PlotHistSum<-function(rep.df,cond.df, y_var, title){
ggplot()+
    geom_jitter(data = HistReplicateSummary,aes(x=as.factor(LogBin),y=.data[[y_var]],  color =Condition ),size=3, height = 0, width = .2, shape = 1)+
    geom_point(data = HistConditionSummary,aes(x=as.factor(LogBin),y=.data[[y_var]],  color =Condition,group = Condition),size=4, shape = 16, position = position_dodge(width = .5) )+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(Species~TimePoint)+
    ggtitle(title)+
    labs(y = paste(y_var,"(%)"),x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    scale_color_brewer(palette = "Set1")+
    theme_cowplot()+
    panel_border() +
    theme(plot.title = element_text(hjust = 0.5), 
        strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facet text
        strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))#facet background
}

@@

#####
#now only aggregates 5 microns or larger
#####
#Separating the replicates
aggsHistRep<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% filter(!LogBin == 5) %>% group_by(Species,Condition,TimePoint, Replicate)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>% group_by(Species,Condition,TimePoint, Replicate, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

##Summary df
aggsHistRepSum<-aggsHistRep %>% group_by(Species,Condition,TimePoint, Replicate, LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint, Replicate) %>% mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))

#Pooling replicates together
aggsHistCond<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% filter(!LogBin == 5) %>% group_by(Species,Condition,TimePoint)%>% 
  mutate(TotalBiomass=sum(Size),TotalFrequency=n()) %>% ungroup() %>% group_by(Species,Condition,TimePoint, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

##Summary df
aggsHistCondSum<-aggsHistCond %>% group_by(Species,Condition,TimePoint, Replicate, LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint, Replicate) %>% mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))%>%ungroup()

@@

#####
#Planktonic vs aggregates
#####
#Add agregation value to the dataframe
##Normalize using replicates
HistReplicate<-HistReplicate %>%  mutate(Aggregation = case_when(LogBin == 5 ~"Planktonic",LogBin>5 ~ "Aggregate")) %>%
  group_by(Species,Condition,TimePoint, Replicate, Aggregation)%>% 
  mutate(AgPlBiomass = sum(Size), AgPlFreq = n()) %>% ungroup() %>%
  mutate(AgPlNormFrequency = AgPlFreq/TotalFrequency*100, AgPlNormBiomass = AgPlBiomass/TotalBiomass*100)
#Collapse to only values per replicate
plankAggHistRep<-HistReplicate %>% group_by(Species,Condition,TimePoint, Replicate, Aggregation,AgPlNormFrequency,AgPlNormBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint, Replicate) %>% mutate(checkAggFreq =sum(AgPlNormFrequency), checkAggBiomass = sum(AgPlNormBiomass))

##Normalize using Condition
HistCondition<-HistCondition %>%  mutate(Aggregation = case_when(LogBin == 5 ~"Planktonic",LogBin>5 ~ "Aggregate")) %>%
  group_by(Species,Condition,TimePoint, Aggregation)%>% 
  mutate(AgPlBiomass = sum(Size), AgPlFreq = n()) %>% ungroup() %>%
  mutate(AgPlNormFrequency = AgPlFreq/TotalFrequency*100, AgPlNormBiomass = AgPlBiomass/TotalBiomass*100)
#Collapse to only values per replicate
plankAggHistCond<-HistCondition %>% group_by(Species,Condition,TimePoint, Aggregation,AgPlNormFrequency,AgPlNormBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint) %>% mutate(checkAggFreq =sum(AgPlNormFrequency), checkAggBiomass = sum(AgPlNormBiomass))

@@

```{r Figures from Pasa SCFM2 Paper}
#Biomass contribution of Planktonic vs Aggregates
#Collapse dataframe to values per condition
plankAggHistSummary<-plankAggHistRep %>% group_by(Species,TimePoint, Condition,Aggregation) %>%

>>>>

```{r aggregate input}
#agregate input functions
ReadAggregateFile<-function(file, species, timepoint, condition, replicate, sampleCode){#make a dataframe of the aggregate size list and add experimental parameters
  print(file)
   datos <-read.csv(file, header = F)  
   colnames(datos)<-c('Size') 
   datos<-datos %>% filter(!Size<0.5) #first filter the size
   datos<-datos %>% mutate(Species = species)#add parameters
   datos<-datos %>% mutate(TimePoint = timepoint) 
   datos<-datos %>% mutate(Condition = condition)
   datos<-datos %>% mutate(Replicate = replicate)
   datos<-datos %>% mutate(SampleCode = sampleCode) 
   datos$Species<-as.factor(datos$Species)
   datos$Replicate<-as.factor(datos$Replicate)
   datos$Condition<-as.factor(datos$Condition)
   datos$TimePoint<-as.factor(datos$TimePoint)
   datos$SampleCode<-as.factor(datos$SampleCode)
   return(datos)
}
GetAggregates<-function(){#import all experiments at once
   folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Aggregate lists')
   fileList<-list.files(path = folder, full.names = TRUE)
   growthDf<-data.frame()
   for (aggList in 1:length(fileList)){
     fileName<-str_split(basename(fileList[aggList]), "\\.")[[1]][1]#this selects the base file name and removes the ".csv"
     attrList<-str_split(fileName,"_")[[1]]
     growthDf<-rbind(growthDf,ReadAggregateFile(fileList[aggList],attrList[4],attrList[2],attrList[1],attrList[3], fileName))
     }
   return(growthDf)
}

#run functions
aggregateList<-GetAggregates()#this reads the data from Matlab
#remove bad samples and irrelevant conditions
##first remove irrelevant conditions
removable <- bind_rows(aggregateList %>% filter(Species == "Pa" & Condition == "mono"), #Pa in Sa mono culture
                       aggregateList%>% filter(Species == "Sa" & Condition == "pawt"),#Sa in all Pa mono cultures
                       aggregateList%>% filter(Species == "Sa" & Condition == "papqsL"),
                       aggregateList%>% filter(Species == "Sa" & Condition == "paphz"))
aggregateList <- anti_join(aggregateList, removable)
##then, bad samples
aggregateList<-aggregateList %>% filter(!SampleCode == "mono_d1_03_Pa" & !SampleCode == "mono_d1_03_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "phz_d4_14_Pa" & !SampleCode == "phz_d4_14_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "pqsL_d4_02_Pa" & !SampleCode == "pqsL_d4_02_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "pqsL_d4_04_Pa" & !SampleCode == "pqsL_d4_04_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "pqsL_d4_10_Pa" & !SampleCode == "pqsL_d4_10_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "wt_d4_07_Pa" & !SampleCode == "wt_d4_07_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "papqsL_d4_02_Pa" & !SampleCode == "papqsL_d4_02_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "pawt_d4_05_Pa" & !SampleCode == "pawt_d4_05_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "phz_d4_06_Pa" & !SampleCode == "phz_d4_06_Sa")

#add proper labels for plots
aggregateList$Species <- factor(aggregateList$Species, levels = c("Pa", "Sa"),labels = c('P. aeruginosa','S. aureus'))
aggregateList$Condition <- factor(aggregateList$Condition, levels = c("mono", "pawt","papqsL", "paphz", "wt", "pqsL", "phz"),
                                  labels = c('Sa mono', 'Pa wt mono','pqsL mono','phz mono','wt co', 'pqsL co', 'phz co'))
aggregateList$TimePoint <- factor(aggregateList$TimePoint, levels = c("d1", "d4"),labels = c('Day 1','Day 4'))

aggregateList<-aggregateList %>% mutate(Culture = recode(Condition, 'Sa mono' = "mono", 'Pa wt mono' = "mono",'pqsL mono'= "mono",'phz mono'= "mono",'wt co'= "co", 'pqsL co'= "co", 'phz co'= "co"))
aggregateList<-aggregateList%>% mutate(Strain = recode(Condition, 'Sa mono' = "Sa", 'Pa wt mono' = "wt",'pqsL mono'= "pqsL",'phz mono'= "phz",'wt co'= "wt", 'pqsL co'= "pqsL", 'phz co'= "phz"))

@@

#Get aggregate summary per sample
aggregateListSummaryReplicate<-aggregateList%>% group_by(Species, Strain, Culture,TimePoint, Condition, Replicate) %>% 
  summarise(Aggs = n(), biomass = sum(Size), meanAggSize = mean(Size), seAggSize = std.error(Size), maxSize = max(Size), minSize = min(Size)) %>%
  ungroup() %>% mutate(RatioMaxTotal = maxSize/biomass )

@@

ggplot()+ 
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= meanAggSize, color = Strain,shape = "MeanSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= maxSize, color = Strain, shape = "MaxSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= biomass, color = Strain, shape = "TotalBiomass"),size = 3, height = 0,width= .3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(Species~Culture, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per sample')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
  theme_cowplot()+
  panel_border() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facet text
        strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))#facet background

@@

#Get aggregate Summary per condition
aggregateListSummary<-aggregateList %>% group_by(Species, Strain, Culture,TimePoint, Condition) %>% 
  summarise(Aggs = n(), biomass = sum(Size), meanAggSize = mean(Size), seAggSize = std.error(Size), maxSize = max(Size), minSize = min(Size)) %>%
  ungroup() %>% mutate(RatioMaxTotal = maxSize/biomass ) 
#plot agg summary per condition
#####
#Version 1
ggplot()+ 
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= meanAggSize, color = Condition,shape = "MeanSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= maxSize, color = Condition, shape = "MaxSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= biomass, color = Condition, shape = "TotalBiomass"),size = 3, height = 0,width= .3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(. ~ Species, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per condition')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5),strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))
#Version 2
ggplot()+ 
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= meanAggSize, color = Strain,shape = "MeanSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= maxSize, color = Strain, shape = "MaxSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= biomass, color = Strain, shape = "TotalBiomass"),size = 3, height = 0,width= .3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(Species~Culture, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per condition')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5),strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))

#####
```

@@

#There are two ways to normalize the Bin Frequncy and Bin Biomass, per condition and per sample, so I'll do both
#First per replicate:
HistReplicate<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% group_by(Species,Culture,Strain, Condition,TimePoint, Replicate)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>% 
  group_by(Species,Culture,Strain, Condition,TimePoint, Replicate, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

#This is weird, but I want to have the histogram above with all the data, and the one below with only the bins
HistReplicateSummary<-HistReplicate %>% 
  group_by(Species,Culture,Strain, Condition,TimePoint, Replicate, LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>%
  unique()%>%ungroup() %>%  group_by(Species,Culture,Condition,TimePoint, Replicate) %>% 
  mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))

#Now per condition. The difference is int the "group_by" part
HistCondition<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% group_by(Species,Culture,Strain, Condition,TimePoint)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>%
  group_by(Species,Culture,Strain, Condition,TimePoint,LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

#Summary data frame
HistConditionSummary<-HistCondition %>%
  group_by(Species,Culture,Strain, Condition,TimePoint,LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>% 
  unique()%>%ungroup() %>%  group_by(Species,Culture,Condition,TimePoint) %>% 
  mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))

#Plot Normalized Frequency
#this isn't useful as the normalized Frequency changes with replicate, so it's better to pool all samples together
PlotHistSum<-function(rep.df,cond.df, y_var, title){
ggplot()+
    geom_jitter(data = rep.df,aes(x=as.factor(LogBin),y=.data[[y_var]],  color =Condition),size=3, height = 0, width = .2, shape = 1)+
    geom_point(data = cond.df,aes(x=as.factor(LogBin),y=.data[[y_var]],  color =Condition,group = Condition),size=4, shape = 16, position = position_dodge(width = .5) )+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(Species~TimePoint)+
    ggtitle(title)+
    labs(y = paste(y_var,"(%)"),x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    scale_color_brewer(palette = "Set1")+
    theme_cowplot()+
    panel_border() +
    theme(plot.title = element_text(hjust = 0.5), 
        strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facet text
        strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))#facet background
}

@@

#####
#now only aggregates 5 microns or larger
#####
#Separating the replicates
aggsHistRep<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% filter(!LogBin == 5) %>% group_by(Species,Culture,Strain,Condition,TimePoint, Replicate)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>% group_by(Species,Culture,Strain,Condition,TimePoint, Replicate, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

##Summary df
aggsHistRepSum<-aggsHistRep %>% group_by(Species,Culture,Strain,Condition,TimePoint, Replicate, LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Culture,Strain,Condition,TimePoint, Replicate) %>% mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))

#Pooling replicates together
aggsHistCond<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% filter(!LogBin == 5) %>% group_by(Species,Culture,Strain,Condition,TimePoint)%>% 
  mutate(TotalBiomass=sum(Size),TotalFrequency=n()) %>% ungroup() %>% group_by(Species,Culture,Strain,Condition,TimePoint, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

##Summary df
aggsHistCondSum<-aggsHistCond %>% group_by(Species,Culture,Strain,Condition,TimePoint, Replicate, LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Culture,Strain,Condition,TimePoint, Replicate) %>% mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))%>%ungroup()

@@

#####
#Planktonic vs aggregates
#####
#Add agregation value to the dataframe 
##Normalize using replicates
HistReplicate<-HistReplicate %>%  mutate(Aggregation = case_when(LogBin == 5 ~"Planktonic",LogBin>5 ~ "Aggregate")) %>%
  group_by(Species,Culture,Condition,Strain,TimePoint, Replicate, Aggregation)%>% 
  mutate(AgPlBiomass = sum(Size), AgPlFreq = n()) %>% ungroup() %>%
  mutate(AgPlNormFrequency = AgPlFreq/TotalFrequency*100, AgPlNormBiomass = AgPlBiomass/TotalBiomass*100)
#Collapse to only values per replicate
plankAggHistRep<-HistReplicate %>% group_by(Species,Culture,Condition,Strain,TimePoint, Replicate, Aggregation,AgPlNormFrequency,AgPlNormBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Culture,Strain,Condition,TimePoint, Replicate) %>% mutate(checkAggFreq =sum(AgPlNormFrequency), checkAggBiomass = sum(AgPlNormBiomass))

##Normalize using Condition
HistCondition<-HistCondition %>%  mutate(Aggregation = case_when(LogBin == 5 ~"Planktonic",LogBin>5 ~ "Aggregate")) %>%
  group_by(Species,Culture,Condition,Strain, TimePoint, Aggregation)%>% 
  mutate(AgPlBiomass = sum(Size), AgPlFreq = n()) %>% ungroup() %>%
  mutate(AgPlNormFrequency = AgPlFreq/TotalFrequency*100, AgPlNormBiomass = AgPlBiomass/TotalBiomass*100)
#Collapse to only values per replicate
plankAggHistCond<-HistCondition %>% group_by(Species,Culture,Condition,Strain,TimePoint, Aggregation,AgPlNormFrequency,AgPlNormBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Culture,Strain, Condition,TimePoint) %>% mutate(checkAggFreq =sum(AgPlNormFrequency), checkAggBiomass = sum(AgPlNormBiomass))

@@

```{r Figures from Pasa SCFM2 Paper}
#Biomass contribution of Planktonic vs Aggregates
#####
#Do stats on the df that has each replicate
plankAggHistSummary<-plankAggHistRep %>% group_by(Species,TimePoint,Culture,Strain,Condition,Aggregation) %>%
  summarise(MeanFreq = mean(AgPlNormFrequency), SEFreq = std.error(AgPlNormFrequency), MeanBiomass = mean(AgPlNormBiomass),
            SEBiomass = std.error(AgPlNormBiomass))
